---
title: "Hosting Rmds with GitHub Pages"
author: "Adam Spannbauer"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    toc_float: true
    includes:
      in_header: "favicon.html"
---

### Preface

This is just 1 way to host an Rmd document.  There are other options.  This is just intended to show how to do it with GitHub pages, it is not claiming to GitHub pages to be the best route for all cases.

# Step-by-step instructions

* [Make a Github repo](https://docs.github.com/en/get-started/quickstart/create-a-repo#create-a-repository)
* Make an R Markdown file you're proud of and want to share with the world!
* [Knit the .Rmd file to and html document](https://rmarkdown.rstudio.com/articles_intro.html#:~:text=To%20transform%20your%20markdown%20file,in%20the%20new%20file%20format.)
* Name the html document `index.html`
* Get the `index.html` file into your GitHub repo
  * Long term I recommend to find a Git+GitHub tutorial and learn how to interact with git.  It's such a useful skill for so many different things!
    * [here's a great one if you're just getting started with git](https://www.youtube.com/watch?v=BCQHnlnPusY&list=PLRqwX-V7Uu6ZF9C0YMKuns9sLDzK6zoiV)
  * Short-term you can use the <kbd>Add file</kbd> button followed by clicking <kbd>Upload files</kbd>

* Turn on GitHub pages for the repo by clicking:
  * <kbd>Settings</kbd>
  * <kbd>Pages</kbd> (under the Code and automation section in the sidebar)
  * Find the dropdown menu in the **Branch** section of the page and select "main" (there's a chance you want to select "master")
    * If you're new to GitHub there's probably only one option to choose, choose that one option that's not "None"
  * Click the <kbd>Save</kbd> button
* Wait!
* Find your website online at:`https://<your-username>.github.io/<your-repo-name/`
  * For example this repo is at https://github.com/AdamSpannbauer/hosting-Rmds-with-GitHub-Pages/
    * This shows my user name is **`AdamSpannbauer`**
    * This shows my repo name is **`hosting-Rmds-with-GitHub-Pages`**
  * It's published GitHub pages site can is https://adamspannbauer.github.io/hosting-Rmds-with-GitHub-Pages/

# Embed other pages/projects in an Rmd

## Embed arbitrary websites

Embed other websites with `knitr::include_url()`!  Shiny's an example option where you might have work to show off.

```{r out.width="100%"}
knitr::include_url("https://spannbaueradam.shinyapps.io/sql-racer/", height = "512px")
```

Using `{r out.width="100%", echo=FALSE}` will set the embed to 100% of its container's width and hide the code that created it.  Eg, the you don't see the embed code for the below example because of `echo=FALSE`.

```{r out.width="100%", echo=FALSE}
knitr::include_url("https://spannbaueradam.shinyapps.io/sp23bas475attendance/", height = "400px")
```

## Embed Tableau Public

* Find you dashboard on [Tableau Public](https://public.tableau.com/app/discover)
* Click the share button
* Click the Copy Link button
* Paste link into an Rmd chunk using `knitr::include_url()`
  * Your link should start with `https://public.tableau.com/views`; if it mentions `user` it will fail (this bullet will prolly fall out of date)
* Update your link so that it shows `?:showVizHome=no&:embed=true` right after the dashboard name
  * My original URL was: `https://public.tableau.com/views/recreate-test-f23/Dashboard1?:language=en-US&:display_count=n&:origin=viz_share_link`
  * My usable URL is: `https://public.tableau.com/views/recreate-test-f23/Dashboard1?:showVizHome=no&:embed=true`

The below code is what creates the embedded view.  If you don't want the code visible use `echo=FALSE` (that's what made the second embedded view here).

Note: This dashboard wasn't designed to be so cramped. If it's for an embedded portfolio piece, take some time to readjust things to work nicely.

```{r out.width="100%", echo=FALSE}
knitr::include_url("https://public.tableau.com/views/recreate-test-f23/Dashboard1?:showVizHome=no&:embed=true", height = "400px")
```

The above was created with:

```
```{r out.width="100%", echo=FALSE}
knitr::include_url("https://public.tableau.com/views/recreate-test-f23/Dashboard1?:showVizHome=no&:embed=true", height = "400px")
```
```

# Make your Rmd more fun!

## Outside resources

I cover some things you can do here, but Rmd is a very flexible thing! Pretty much anything you can do in R Markdown could be applied here. See the below online free resources for a very complete overview of what can be done:

* [*R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/)
* [*R Markdown Cookbook*](https://bookdown.org/yihui/rmarkdown-cookbook/html-css.html)

## What the h*ck is a YAML?

A lot of what's covered below deals with the "[YAML front matter](https://assemble.io/docs/YAML-front-matter.html)" that appears at the beginning of an .Rmd file.

The YAML used for this document is:

```yaml
---
title: "Hosting Rmds with GitHub Pages"
author: "Adam Spannbauer"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    toc_float: true
    includes:
      in_header: "favicon.html"
---
```

All the contents of the the YAML are either self-explanatory (message me if not!) or they're discussed below. The exception is the usage of `Sys.Date()` in the `date:` field; this will tell the .Rmd file to look up today's date when you click the knit button (ie the date will serve as a "last updated" date).

## CSS themes

Adding a theme is easy to do and can make documents look snazzy and professional.

### Easy mode (what I do)

Use the below YAML to tell your .Rmd what theme to use.

```yaml
output: 
  html_document:
    theme: flatly
```

The options you can say other than `flatly` used above are: cerulean, cosmo, cyborg, darkly, flatly, journal, lumen, paper, readable, sandstone, simplex, slate, spacelab, superhero, united, yeti. So many!!!

### CSS knowers and lovers

If you're already in the know about CSS, it's worth noting that you can apply you own style sheet to an Rmd file.  See the [official R Markdown Cookbook here](https://bookdown.org/yihui/rmarkdown-cookbook/html-css.html).

This can be a great option if you really want to work towards unique consistent branding.

## Table of contents

Again, a table of contents is just a little piece of YAML away.  Below is what this doc uses:

```yaml
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
```

* `toc: true` - turn on the table of contents
* `toc_depth: 2` - the table of contents should include both first and second level headers (ie sections and sub-sections, not sub-sub-sections etc)
* `toc_float: true` - the table of contents should always appear on the side. Set to `false` and the table of contents will be stuck to the top of the page (ie when users scroll down it will scroll off the top of the page)

See [this section from *R Markdown: The Definitive Guide*](https://bookdown.org/yihui/rmarkdown/html-document.html#table-of-contents) for more options for the table of contents!

## Favicon

A little icon in the tab of your web browser can go a surprisingly long way in making a site look more professional. This little icon in the browser tab is known as a "favicon".  To add one, I've included the below in my YAML:

```yaml
output: 
  html_document:
    includes:
      in_header: "favicon.html"
```

, and then I have [this small html file](https://raw.githubusercontent.com/AdamSpannbauer/hosting-Rmds-with-GitHub-Pages/master/favicon.html) saved as `favicon.html` in the same folder as my .Rmd file.  To change the icon from the little emoji R, you'd change the link to an image you'd like to be a favicon.  They are intended to be small images, so there's a chance you don't use this feature perfectly as intended by web browsers.

## Interactive graphics

```{r message=FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
```

Part of the fun of being on the web instead of in a Word or PDF doc is that you can use interactive things in your Rmd!

For example here's a plot made with `ggplot2` and then rendered with `plotly` to add some nice interactivity.

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    x = "Car weight in 1000s of lbs",
    y = "MPG"
  )

ggplotly(p)
```

You can also use `plotly` directly to do some pretty cool things.

Data viz note: Is this the best way to show this data? Prolly not, as always double check the story you want to tell and focus on that. But animations are fun to play with!

```{r message=FALSE, warning=FALSE}
keep_cities <- c(
  "Arlington", "Austin", "Dallas",
  "El Paso", "Fort Worth", "Houston",
  "Lubbock", "San Antonio"
)

annual_txhousing <- txhousing %>%
  filter(city %in% keep_cities) %>%
  group_by(city, year) %>%
  summarise(
    sales = sum(sales),
    volume = sum(volume)
  )

plot_ly(
  annual_txhousing,
  x = ~ log(sales),
  y = ~ log(volume),
  color = ~city,
  frame = ~year,
  type = "scatter",
  mode = "markers"
) %>%
  layout(
    xaxis = list(title = "log(Number of Sales)"),
    yaxis = list(title = "log(Total value of Sales)")
  )
```
